Catchy Title Goes Here

### 2018-05-22 Paper/paper.tex
Updated introduction
Added introduction

\section{Introduction \label{sec:Intro}}
- Quantifying the size and age distribution of the population susceptible to measles is a critical tool in evaluating vaccination programs and developing interventions (Winter et al JID; Cutts and Hanson TMIH).  
- Unfortunately, direct observation of the susceptible population is challenging as it requires a sero-survey which can be cost-prohibitive (Winter et al JID paper). 
- Previous efforts have estimated the susceptible distribution though demographic models that account for inputs through births and immunization via both vaccination an natural infection.
	- Winter et al (new paper on Madagascar)
	- Merler and the Italian group (http://www.thelancet.com/pdfs/journals/laninf/PIIS1473-3099(17)30421-8.pdf)
- The contribution of natural infection to population immunity can be challenging to estimate. In particular, the age-specific force of infection is not directly estimable from aggregate population time series ? (both papers referenced above rely on SIR models that assume a specific age-structured mixing). 
	- challenges of using historical incidence data -- under-reporting, changing reporting.

- The catalytic model has previously been used to estimate the age-specific force of infection from both serological surveys and the cross-sectional age-distribution of measles cases.  The catalytic model, fit to the cross-sectional age distribution of cases necessarily returns an estimate of the age distribution of susceptibles as a by-product.  This suggests that routine case surveillance could be used to generate an estimate of the age distribution of susceptibles.

- Competing rates ? 
In the vaccine era, however, the age distribution of susceptibility is generated by the sum of the rates of vaccination and natural infection.

Original applications of the catalytic model assumed that natural infection was the only source of immunity.

Subsequent analyses have made simple assumptions about the role of vaccination, in particular that the delivery of vaccination is timely, and thus affects the age distribution of the susceptible population only by reducing the force of natural infection.  However, recent work has highlighted significant variability both in the maximum vaccination coverage achieved and the timeliness of vaccination ? that is, the proportion of children receiving vaccination after the recommended age.  Thus, failing to account for the age-specific pattern of vaccination may lead to a biased interpretation of case-data, as children who might be vaccinated later than the recommended age remain susceptible may contribute to incident cases.

@@

Routine vaccination coverage surveys and age-specific case reporting are more readily available data sources (see Takahashi text re: DHS).

Lay out the DRC case study?

Here we present a novel formulation of the catalytic model to estimate the age-specific sero-prevalence, and also the age-specific rates of vaccination and force of infection, and vaccine effectiveness using age-specific vaccine coverage data and case-records. We illustrate performance of this model using both simulated data, and measles case surveillance data from DRC combined with vaccination coverage surveys conducted as part of the 2013-14 DHS.  A contemporary measles sero-survey conducted during the 2013-14 DHS allows us to validate the performance of our estimates of sero-prevelence against direct measurements. We finally present a fit of the model to the surveillance data, vaccine coverage data, and sero-survey and discuss opportunities for combining data sources ?

>>>>

\section{Introduction \label{sec:Intro}}
Dramatic improvements in the coverage of vaccination with measles containing vaccines have lead to significant reductions in the incidence of measles disease and thea associated childhood mortality [REF - Dabbagh et al]. 
However, despite these improvements at the national scale, significant heterogeneity in both vaccination coverage and disease incidence remains at the sub-national level [REF Takahashi].
Given the high transmission rate of measles, the proportion of the population that must be immunized to achieve and maintain elimination of measles is likely to be greater than 90\% in most populations [REF - Goodson?] (the exact level will be modulated by locally specific conditions just as popualtion density and contact rates).
%"The WHO recommends that programs use good quality data to monitor population immunity by identifying and responding to large numbers of susceptibles" from Winter et al reference to WHO Measles vaccines: WHO position paper Wkly Epidemiol Rec 2017;92:205-28
%"Because the risk of measles outbreaks is determined by the rate of accumulation of susceptible people in the population, programmes should use available good quality data on population immunity (i.e. vacci- nation coverage, surveillance, serological studies) to monitor the accumulation of susceptible people and conduct follow-up campaigns before the number of pre-school children susceptible to measles approaches the equivalent of one birth cohort, in order to prevent an outbreak of measles. This approach has been found to be programmatically useful in preventing large outbreaks. However, for large countries and countries which are close to measles elimination, a more exten- sive assessment of the accumulation of susceptible persons should be carried out at the subnational level." WHO Measles vaccines: WHO position paper Wkly Epidemiol Rec 2017;92:205-28 -- http://apps.who.int/iris/bitstream/handle/10665/255149/WER9217.pdf?sequence=1

The size and age-distribution of the population susceptible to measles (hereafter we refer to this as the "susceptible persons") depends on the competing prossess of susceptible recruitment through birth and susceptible loss through either natural infection or vaccination. 
Quantifying the size and age distribution of susceptible persons to measles is a critical tool in evaluating outbreak risk, the performance of vaccination programs, and developing vaccine-based interventions. 
The WHO recommends that programs monitor the accumulation susceptible persons using "good quality data" for all countries and that these should be made at the sub-national level for large countries and those close to measles elimination [REF -- Wkly Epid Rec].

Direct observation of the distribution of susceptible persons is challenging as it requires a sero-survey which can be cost-prohibitive (Winter et al JID paper). 
A sero-survey can provide a well resolved cross-sectional estimate of the distribution of suscpetible persons, it cannot, by itself, quantify the relative contribution of natural and vaccine derived immunity as current serological diagnostics cannot distinguish between these two sources [REF - Winter JID]; however, the relative contribution of these sources can be estimated through modeling [REF- Winter JID].
In the absence of a sero-survey, the number and age-distribution of susceptible persons can be estimated though demographic models that account for inputs through births and immunization via both vaccination an natural infection [Winter et al JID, Winter et al Madagascar, Merler et al http://www.thelancet.com/pdfs/journals/laninf/PIIS1473-3099(17)30421-8.pdf]. In this approach, the contribution of natural immunity is modeled as the result of a dynamic transmission model [REF] which requires explicit assumptions about the age-sepcific force of infection via a Who Acquires Infection From Whom matrix that describes the age-specific rate of infectious contacts from each age class to each other.  Direct estimation of this WAIFW matrix is challenging from disease incidence or seroprevalence data [though see Whittaker and Farrington etc.], and thus it is common to make simplfying assumptions that either the WAIFW matrix is constant for all ages (i.e. a uniformly well-mixed system) or that the WAIFW matrix is a scalar function of some other measureable interaction process, such as contacts measures via diary studies [REF -- Mossong POLYMOD study].

The catalytic model was initially proposed as a method for estimating the age-specific force of infection from cross-sectional age-specfic sero-prevalence data [REF - Griffiths]. 
This model, which represents the probability of immunization via natural infection at specific age $a$ as the cumulative sum of age-specific force of infection prior to age $a$, was then adapted by Grenfell and Anderson [REF - 1985] for use with cross-sectional, age-specific case observations.  
The resulting fitted model can then be used to estimate the age-specific distribution of susceptible persons.
Thus, in the absence of a sero-survey it is possible to estimate seroprevalence from routinely collected age-specific case reporting.

Original applications of the catalytic model assumed that natural infection was the only source of immunity [REF - Griffiths].  
In the vaccine era, however, the age distribution of susceptibility is generated by the sum of the rates of vaccination and natural infection. 
Though measles vaccination is recommended to follow a specific schedule, with children receiving two doses at prescribed ages, in practice, the ages at which receive an immunizing dose of measles containing vaccine (MCV) may be highly variable because of variaion in access to vaccination services [REF Metcalf; Takahashi; others] and the multiple vaccination initiatives that are employed: e.g. routine vaccination, supplementary immunization activities, child and maternal health days, school-based immunization drives, etc.
Recent work has highlighted significant variability both in the maximum vaccination coverage achieved and the timeliness of vaccination, that is, the proportion of children receiving vaccination at any given age.
Thus, vaccination can itself be modeled as an age-specific hazard rate that accounts for the sum of multiple forces.
Failing to account for the age-specific pattern of vaccination may lead to a biased interpretation of case-data, as children who might be vaccinated later than the recommended age remain susceptible may contribute to incident cases.

@@

Here we present an extension of the catalytic model that represents the age-specific proportion immune as the cumulative sum of the hazards of both natural infection (force of infection) and effective immunization via vaccination.
We illustrate how this model can be used to estimate the age-specific sero-prevalence, and also the age-specific rates of vaccination and force of infection, and vaccine effectiveness using age-specific vaccine coverage data and case-records. 
We illustrate performance of this model using both simulated data, and measles case surveillance data from DRC combined with vaccination coverage surveys conducted as part of the 2013-14 DHS.  
A contemporary measles sero-survey conducted during the 2013-14 DHS allows us to validate the performance of our estimates of sero-prevelence against direct measurements. 
We finally present a fit of the model to the surveillance data, vaccine coverage data, and sero-survey and discuss opportunities for combining data sources.
